A Discrepancy-Based Algorithm for Extended Inversion

### 2020-07-10 repro/papers/discrep/paper.tex
interim

\title{A Discrepancy-Based Algorithm for Extended Inversion}
\author{William. W. Symes \thanks{P. O. Box 43, Orcas WA USA, email {\tt symes@rice.edu}.}}

@@

FIeld recording takes place over finite time intervals (and with
finite sample rate, a detail I shall ignore in this paper). Denote by
$[t_{\rm min},t_{\rm max}]$ the recording interval for the
hypothetical single trace data. Then natural choices for domain and
range of $F[m]$ are $W = L^2(\bR)$ and $D=L^2([t_{\rm min},t_{\rm
  max}])$ respectively (that is, interpret the definition \ref{eqn:mod}
as holding for $t \in [t_{\rm min},t_{\rm  max}]$).

@@

The support constraint is closely linked to the folk theorem about FWI noted many
times in the literature: convergence of a descent method requires that
the initial slowness must be known to ``within a (fraction of a) wavelength''. The
relation is a consequence of Heisenberg's inequality
(see for example \cite{Folland:07}, p. 255). Assume that $w \in L^2(\bR)$, and that both $tw$ and derivative $dw/dt$ are
square-integrable (an essential additional assumption, as will be
explained below). Then for any $t_0, k_0 \in \bR$,
\[
  \left(\int\,dt\, (t-t_0)^2|w(t)|^2\right)
  \left(\int\,dk\,(k-k_0)^2|\hat{w}(k)|^2\right)
  \ge\frac{\|w\|^4}{16 \pi^2}
\]
in which $\hat{w}$ denotes the Fourier transform of $w$. Rearrange
slightly and choosing $t_0=0, k_0=0$,
\[
  \left(\int\,dt\, t^2\frac{|w(t)|^2}{\|w\|^2}\right)
  \left(\int\,dk\,k^2\frac{|\hat{w}(k)|^2}{\|\hat{w}\|^2}\right)
  \ge\frac{1}{16 \pi^2}
\]
If $w \in \lW$, then the first factor is $\le \lambda^2$, whereas the
second factor is the power spectrum mean square estimate of
frequency squared $k^2$. That is the power spectrum RMS estimate of
frequency is at least $1/(4 \pi \lambda)$. In the sense of the RMS
spectral estimator of frequency, then, $\lambda$ is proportional to an upper
bound on representative wavelength: substantial energy must be present
at frequencies larger than $1/\lambda$.

@@

At this point I have to come clean about the actual choice of
$A$. The role of $A$ is to penalize distance from the constrained
subspace $\lW$, in some sense.

I will introduce two options for $A$. Both are scalar 
multiplication operators defined by a choice of $a \in L^{\infty}(\bR)$:
\begin{equation}
  \label{eqn:annmult}
  A w(t)= a(t)w(t), \, t\in \bR.
\end{equation}

An
operator $A_{\lambda}$ of which $\lW$ is the
null space would be a natural 
choice. Such operators have come to be called ``annihilators'', since
they map all members of the constraint subspace $\lW$ to zero.
Define 
\begin{eqnarray}
  A = E^c_{\lambda}&=&I - E_{\lambda},\mbox{ where } \nonumber \\
  E_{\lambda}w(t) &=&{\bf 1}_{[-\lambda,\lambda]}(t)w(t). 
                      \label{eqn:ann0}
\end{eqnarray}
That is, $E_{\lambda}$ is the orthogonal projector onto $\lW$,
and $E_{\lambda}^c$ is the orthogonal projector onto its
orthocomplement, an operator of the form \ref{eqn:annmult} with $a
= 1 - {\bf 1}_{[-\lambda,\lambda]}$.

A second possible
penalty operator penalizes energy away from
$t=0$: choose $\tau > 0$ and set
\begin{equation}
  \label{eqn:ann}
  a(t) = \min(|t|, \tau).
\end{equation}
The cutoff $\tau$ will be chosen large enough to be effectively inactive:
specifically, hindsight suggests
\begin{equation}
  \label{eqn:taudef}
  \tau = \max \{t_{\rm min}-m_{\rm max}r, t_{\rm max}-m_{\rm min}r\}.
\end{equation}
This
particular annihilator has been employed in earlier papers on extended
source inversion
\cite[]{Plessix:00a,LuoSava:11,Warner:14,HuangSymes:SEG15a,Warner:16,HuangSymes:GEO17}.

@@

\begin{theorem}
  \label{thm:norminv}
  With the assumptions \ref{eqn:mod}, \ref{eqn:fullrec},
  \ref{eqn:annmult} and either choice of penalty multiplier $a$
  described above,
  \begin{itemize}
  \item[1. ]The normal operator $F[m]^TF[m] + \alpha^2A^TA$ is
    invertible for any $m \in M$, $\alpha > 0$;
  \item[2. ] For any $d \in D$,
    \begin{equation}
      \aw[m;d](t) = \left\{
        \begin{array}{c}
          \left(\frac{1}{(4\pi r)^2} + \alpha^2
          a^2(t)\right)^{-1}\frac{1}{4 \pi r}d(t+mr), t \in [t_{\rm
          min}-mr, t_{\rm max}-mr];\\
          0, \mbox{ else.}
        \end{array}
      \right.
    \end{equation}
  \item[3. ] If $d$ is noise-free, as in equation
    \ref{eqn:datadefnonoise},
    \begin{equation}
      \label{eqn:solnnon}
      \aw[m,d](t)= \left(1+ (4\pi r)^2\alpha^2 a(t)^2\right)^{-1}w_*\left(t+(m-m_*)r\right) 
    \end{equation}
  \end{itemize}
\end{theorem}

Sincee normal operator is invertible, the reduced
objective $\tJa$ is given by the expression \ref{eqn:objexp}.
Employing the expression \ref{eqn:tran} for the action of $F[m]^T$,
the definition \ref{eqn:defdatanonoise} of noise-free data,
and the realization \ref{eqn:norm2} of the normal operator as a
multiplication operator, obtain after some simplification

\begin{equation}
  \label{eqn:residnorm}
  e[m,\aw[m,d];d] 
= 8 \pi^2 r^2 \alpha^4\int\,dt\,a(t-(m-m_*)r)^4(1+(4\pi r)^2 \alpha^2 
    a(t-(m-m^*)r)^2)^{-2}w_*(t)^2.
\end{equation}
\begin{equation}
  \label{eqn:anninorm}
  p[m,\aw[m,d];d] = \frac{1}{2}\int \,dt\,  a(t-(m-m_*)r)^2 \left(1+ (4\pi r)^2\alpha^2
    a(t-(m-m_*)r)^2\right)^{-2}w_*(t)^2.
\end{equation}
so finally
\begin{equation}
\label{eqn:expj}
\tJa[m;d] = =\frac{\alpha^2}{2}\int\,dt\,a(t-(m-m_*)r)^2(1+(4\pi r)^2 \alpha^2 
  a(t-(m-m^*)r)^2)^{-1}w_*(t)^2.
\end{equation}

\begin{theorem}
  \label{thm:boxcarbad}
  For $A=E^c_{\lambda}$, that is, $a=1-{\bf 1}_{-\lambda,\lambda]}$ in
  the definition \ref{eqn:expj}. Then if $|m-m_*| >  2\lambda/r$,
  \[
    \tJa[m;d] = \frac{\alpha^2}{2(1+(4 \pi r)^2 \alpha^2)}\|w_*\|^2.
  \]
\end{theorem}

@@

Surprisingly, the choice $A=S$ yields an optimization
problem has better global convergence properties:

\begin{theorem}
  \label{thm:rampgood}
  For $A=S$, that is, $a(t)=t$ in the definition \ref{eqn:expj}, all
  stationary points of $\tJa$ lie in the interval
  \[
    [m_*-\lambda/r, m_*+\lambda/r].
  \]
\end{theorem}
\begin{proof}
  A brief calculation yields
\begin{equation}
\label{eqn:gradfinal}
\frac{\partial}{\partial m}\tJa[m;d] = -r\alpha^2\int \,dt\,t\left(\frac{1}{(4\pi r)^2} + \alpha^2 t^2\right)^{-2}\frac{1}{(4\pi r)^2}w_*\left(t+(m-m_*)r\right)^2
\end{equation}
Recall that $w_*(t+(m-m_*)r)$
vanishes if $|t+(m-m_*)r| > \lambda$. Therefore the integral on the
RHS of equation \ref{eqn:gradfinal} can be re-written
\[
  = -r\alpha^2\int_{-(m-m_*)r-\lambda}^{-(m-m_*)r+\lambda}
  \,dt\, \frac{t}{(1+(4\pi r)^2\alpha^2 t^2)^2}w_*\left(t+(m-m_*)r\right)^2
\]
If $m > m_*+\lambda/r$, then the entire interval of integration is a proper
subset of the negative half-axis. Consequently the first factor in the
integrand satisfies
\[
 \frac{t}{1+(4\pi r)^2\alpha^2 t^2}
  \le -(m-m_*)r+\lambda < 0
\]
over the interval of integration. The integral is
negative, so the derivative
is positive. Similar reasoning applies to the case $m <
m_*-\lambda/r$. Thus
\begin{eqnarray}
  m > m_*+\lambda/r & \Rightarrow & \frac{\partial \tJa[m;d]}{\partial
    m}> 0; \nonumber\\
  m < m_*-\lambda/r & \Rightarrow & \frac{\partial \tJa[m;d]}{\partial
                                    m}< 0.
\label{eqn:leftright}
\end{eqnarray}
\end{proof}

\noindent {\bf Remark.} That is, {\em with the choice $A=S$, $\tJa$ has no local minima
  further than  $O(\lambda)$ from the global minimum:} the gradient
has the correct sign and slowness updates computed from it will be
constructive, unless the slowness estimate is already ``within a
wavelength'' of being correct.

\section{Extended Inversion}
The preceding theorem established that a proper choice of annihilator leads to a
reduced penalty objective all of whose stationary points are within
$O(\lambda)$ of the target slowness $m_*$, provided that the data are
noise-free. This result leaves open two questions:
\begin{itemize}
\item how does one use this reduced penalty minimization to produce
  a solution of the inverse problem as in problem statement
  \ref{eqn:probstat0}? There is no reason to think that the minmizer
  $\bar{m}$ of $\tJa$ with produce $\aw[\bar{m};d] \in \lW$, as
  required by the problem statement.
\item how does one answer the same question for noisy data?
\end{itemize}

A partial resolution of these issues lies in the relation between the
annihilator $S$ (multiplication by $|t|$) and the family of projectors
$\{E_{\lambda}: 0 \le \lambda \le \lambda_{\rm max}\}$. In fact, this
family of operators is a resolution of the identity for the positive
semidefinite self-adjoint $S$:
\begin{eqnarray}
  \label{eqn:res1}
  S &=& \int_{0}^{\|A\|} \lambda d\lE,\nonumber\\
  S^2=S^TS  &=& \int_{0}^{\|A\|} \lambda^2 d\lE.
\end{eqnarray}
This relation is similar to that explained in the example on p. 313 of \cite{Yosida}.

An immediate consequence  is that under some
circumstance, establishing a bound on $\Ja[m,w;d]$ delivers a solution
of the inverse problem in the sense of the Problem Statement,
above.  This conclusion follows from a simple fact about quadratic
penalty functions posed in terms of the spectral resolution of the
penalty operator:

\begin{theorem}
  \label{thm:minsol}
  Suppose $W$ and $D$ are Hilbert spaces, $S \in {\cal B}(W,W)$, $R
  \in {\cal B}(W,D)$, $d \in D, w \in W$, and $\alpha, \epsilon>0, \gamma >
  1$. Suppose $\lambda_{\rm min} \le \lambda_{\rm max}$ and that
  $\{\lE: \lambda \in [\lambda_{\rm min},\lambda_{\rm max}]\}$ is the
  spectral resolution of $\sqrt{S^TS}$. Then for any $\lambda \in  [\lambda_{\rm min},\lambda_{\rm max}]$,
  \begin{equation}
    \label{eqn:gentol}
    \|R\lE w-d\| \le \epsilon\|d\|
  \end{equation}
  provided that 
  \begin{equation}
    \label{eqn:epslam}
    \epsilon \ge \gamma \frac{(\|Rw-d\|^2 +
      \alpha^2\|Sw\|^2)^{1/2}}{\|d\|}, \,\,\lambda \ge 
    \frac{\|R\|}{\alpha (\gamma-1)}
  \end{equation}
\end{theorem}
\begin{proof}
First note that
\[
  \|Sw\|^2 = \|S(\lE w + (I-\lE)w)\|^2 =
  \langle \lE w + (I-\lE)w, S^TS (\lE w + (I-\lE)w)\rangle
\]
\[
  =  \langle \lE w + (I-\lE)w, (\lE w + (I-\lE) S^TS w)\rangle
\]
\[
  = \langle \lE w,\lE S^TSw \rangle+ \langle(I-\lE)w,(I-\lE) S^TS w\rangle
\]
(since $\lE$ is a projector)
\[
\ge  \langle (I-\lE)w,S^TS (I-\lE)w \rangle
\]
\[
  =\langle w, S^TS (I-\lE)w\rangle
\]
(since $(I-\lE)$ commutes with $S^TS$)
\[
  =\int_{\lambda}^{\lambda_{\rm max}} \mu^2 d\langle w,E_{\mu}w\rangle 
\]
\[
  \ge \lambda_{\rm min}^2 \int_{\lambda}^{\|A\|} d\langle w,E_{\mu}w\rangle
\]
\begin{equation}
  \label{eqn:lb}
  =\lambda^2\langle w, (I-\lE)w\rangle   =\lambda^2\|(I-\lE)w\|^2
\end{equation}

Set
\[
  \delta =  \frac{(\|Rw-d\|^2+\alpha^2\|Sw\|^2)^{1/2}}{\|d\|}.
\]
For any $\lambda >0$,
\[
  \alpha \lambda \|(I-\lE)w\| \le \alpha \|Sw\| \le \delta \|d\|^2
\]
whence
\[
  \|R\lE w-d\| \le \|Rw-d\|+\|R(I-\lE)w\|
\]
\[
  \le \delta\|d\| + \|R\|\frac{\delta}{\alpha \lambda}\|d\|
\]
\[
  \le \left(1 + \frac{\|R\|}{\alpha\lambda}\right)\delta\|d\|
\]
The presumed bounds \ref{eqn:epslam} on $\epsilon$ and $\lambda$ imply that this is
\[
  \le \gamma \delta \|d\| \le \epsilon\|d\|.
\]
\end{proof}

In the context of the single trace transmission problem, the operators

\begin{corollary}
  Suppose that $m \in M, d\in D, \gamma>1, \epsilon>0, \lambda >0, \alpha
  > 0$ satisfy
    \begin{equation}
    \label{eqn:epslam}
    \epsilon \ge \gamma \frac{(2\tJa[m;d])^{1/2}}{\|d\|}, \,\,\lambda \ge 
    \frac{\|F[m]\|}{\alpha (\gamma-1)}
  \end{equation}
  Then $(m,\lE \aw[m;d])$ is a solution of the inverse problem as stated
  in condition \ref{eqn:probstat0}, that is, $\lE\aw[m;d] \in \lW$ and
  \[
    e[m,\aw[m;d];d] \le \frac{\epsilon^2}{2} \|d\|^2
  \]
\end{corollary}

That is, the minimizer of $\Ja$ is a solution of the constrained
problem, as stated in the Problem Statement \ref{eqn:probstat0}, for
$\lambda = O(1/\alpha)$, provided that the target noise level
($\epsilon$) is higher by a fudge factor $\gamma>1$ than the minimum
value of $\Ja$.

It remains to establish how

>>>>

\title{A Discrepancy-Based Algorithm for Extended Inversion}
\author{William W. Symes, Orcas Island, Huiyi Chen* and Susan E. Minkoff, The University of Texas at Dallas}

@@

FIeld recording takes place over finite time intervals (and with
finite sample rate, a detail I shall ignore in this paper). Denote by
$[t_{\rm min},t_{\rm max}]$ the recording interval for the
hypothetical single trace data. Then natural choices for domain and
range of $F[m]$ are $W = L^2(\bR)$ and $D=L^2([t_{\rm min},t_{\rm
  max}])$ respectively (that is, interpret the definition \ref{eqn:mod}
as holding for $t \in [t_{\rm min},t_{\rm  max}]$). So
\begin{equation}
  \label{eqn:mapprop}
  \mbox{for }m \in M, F[m] \in {\cal B}(W,D),\mbox{ and }\|F[m]\| =
  \frac{1}{4\pi r}.
\end{equation}

@@

\noindent{\bf Remark:} The support constraint is closely linked to the
folk theorem about FWI noted many times in the literature: convergence
of a descent method requires that the initial slowness must be known
to ``within a (fraction of a) wavelength''. The relation is a
consequence of Heisenberg's inequality, and will be reviewed later in
the paper.

@@

At this point I have to come clean about the actual choice of
$A$. The role of $A$ is to penalize distance from the constrained
subspace $\lW$, in some sense. All choices for $A$ considered here are scalar 
multiplication operators defined by a choice of $a \in L^{\infty}(\bR)$:
\begin{equation}
  \label{eqn:annmult}
  A w(t)= a(t)w(t), \, t\in \bR.
\end{equation}

@@

\begin{theorem}
  \label{thm:norminv}
  Assume the conditions \ref{eqn:mod}, \ref{eqn:fullrec},
  \ref{eqn:annmult}. Also assume that $\lambda \in (0,\lambda_{\rm
    max}], \alpha > 0,$ and that $a \in L^{\rm infty}(\bR)$
  mentioned in condition \ref{eqn:annmult} satisfies
  \begin{equation}
    \label{eqn:abnd}
    |t| > \lambda \Rightarrow a(t) \ge C,
  \end{equation}
  in which $C>0$ may depend on $\lambda$.
  Then
  \begin{itemize}
  \item[1. ]the normal operator $F[m]^TF[m] + \alpha^2A^TA$ is
    invertible for any $m \in M$, $\alpha > 0$;
  \item[2. ]for any $d \in D$, the normal equation \ref{eqn:norm} has
    a unique solution $\aw[m;d]\in W$, and 
    \begin{equation}
      \label{eqn:normsol}
      \aw[m;d](t) = \left\{
        \begin{array}{c}
          \left(\frac{1}{(4\pi r)^2} + \alpha^2
          a^2(t)\right)^{-1}\frac{1}{4 \pi r}d(t+mr), t \in [t_{\rm
          min}-mr, t_{\rm max}-mr];\\
          0, \mbox{ else;}
        \end{array}
      \right.
    \end{equation}
  \item[3. ]if in addition $d=F[m_*]w_*, w_* \in \lW$ is noise-free, as in equation
    \ref{eqn:defdatanonoise},
    \begin{equation}
      \label{eqn:solnnongen}
      \aw[m,d](t)= \left(1+ (4\pi r)^2\alpha^2 a(t)^2\right)^{-1}w_*\left(t+(m-m_*)r\right).
    \end{equation}
  \end{itemize}
\end{theorem}

\begin{proof}
  \begin{itemize}
  \item[1. ]Note that thanks to \ref{eqn:zeroinc}, if $|t|\le
    \lambda \le \lambda_{\rm max}$, then ${\bf 1}_{[t_{\rm min}-mr,  
      t_{\rm max}-mr]}(t) = 1$, whereas if $|t|>\lambda$,
    then $a(t) \ge C$, whence
    \[
      \frac{1}{(4\pi r)^2}  {\bf 1}_{[t_{\rm min}-mr,  
        t_{\rm max}-mr]} + \alpha^2a^2  \ge \min\{(4\pi r)^2,
      \alpha^2\min\{1/(4\pi r)^2,C^2\} \}> 0.
    \]
    Therefore the normal operator is invertible under the stated
    conditions.

  \item[2. ]From the identity \ref{eqn:tran}.
    \[
      \mbox{supp }F[m]^Td \subset [t_{\rm min}-mr,t_{\rm max}-mr].
    \]
    Define $w_{\rm tmp}$ to be the right-hand side of equation \ref{eqn:normsol}. Then
    from the previous observation and identity \ref{eqn:tran},
    \[
      \mbox{supp }w_{\rm tmp} \subset [t_{\rm min}-mr,t_{\rm max}-mr].
    \]
    From the identity \ref{eqn:unit}, for any $w \in W$,
    \[
      t \in [t_{\rm min}-mr,t_{\rm max}-mr] \Rightarrow F[m]^TF[m]w(t)
      = \frac{1}{(4 \pi r)^2}w(t).
    \]
    It follows from this and the previous two observations that
    $w_{\rm tmp}$ solves the normal equation \ref{eqn:norm}, and
    therefore that $\aw[m;d]=w_{\rm tmp}$.

  \item[3. ]Follows by inserting the definition
    \ref{eqn:defdatanonoise} of $d$ in \ref{eqn:normsol} and
    rearranging.
  \end{itemize}
\end{proof}

Since the normal operator is invertible, the reduced
objective $\tJa$ is given by the expression \ref{eqn:objexp}.
The results of the previous theorem imply those of

\begin{theorem}
  \label{thm:epjnonoise}
  Assume the hypotheses of Theorem \ref{thm:norminv}, item 3. Then 
\begin{equation}
  \label{eqn:residnorm}
  e[m,\aw[m,d];d] 
= 8 \pi^2 r^2 \alpha^4\int\,dt\,a(t-(m-m_*)r)^4(1+(4\pi r)^2 \alpha^2 
    a(t-(m-m^*)r)^2)^{-2}w_*(t)^2.
\end{equation}
\begin{equation}
  \label{eqn:anninorm}
  p[m,\aw[m,d];d] = \frac{1}{2}\int \,dt\,  a(t-(m-m_*)r)^2 \left(1+ (4\pi r)^2\alpha^2
    a(t-(m-m_*)r)^2\right)^{-2}w_*(t)^2.
\end{equation}
so finally
\begin{equation}
\label{eqn:expjgen}
\tJa[m;d] = =\frac{\alpha^2}{2}\int\,dt\,a(t-(m-m_*)r)^2(1+(4\pi r)^2 \alpha^2 
  a(t-(m-m^*)r)^2)^{-1}w_*(t)^2.
\end{equation}
\end{theorem}

I willl now examine two possible choices of penalty multipler $a$ in
defintion \ref{eqn:annmult}.

A penalty operator $A$ of which $\lW$ is the null space would be a
natural choice. Such operators have come to be called
``annihilators'', since they map all members of the constraint
subspace $\lW$ to zero.  A simple example is
\begin{eqnarray}
  A = E^c_{\lambda}&=&I - E_{\lambda},\mbox{ where } \nonumber \\
  E_{\lambda}w(t) &=&{\bf 1}_{[-\lambda,\lambda]}(t)w(t). 
                      \label{eqn:ann0}
\end{eqnarray}
That is, $E_{\lambda}$ is the orthogonal projector onto $\lW$,
and $E_{\lambda}^c$ is the orthogonal projector onto its
orthocomplement, an operator of the form \ref{eqn:annmult} with $a
= 1 - {\bf 1}_{[-\lambda,\lambda]}$. Note that this choice of $a$
satisfies condition \ref{eqn:abnd} with $C \equiv 1$.

\begin{theorem}
  \label{thm:boxcarbad}
  For $A=E^c_{\lambda}$, that is, $a=1-{\bf 1}_{-\lambda,\lambda]}$ in
  the definition \ref{eqn:expjgen}. Then if $|m-m_*| >  2\lambda/r$,
  \[
    \tJa[m;d] = \frac{\alpha^2}{2(1+(4 \pi r)^2 \alpha^2)}\|w_*\|^2.
  \]
\end{theorem}

@@

A second possible 
penalty operator penalizes energy away from 
$t=0$: choose $\tau > 0$ and set 
\begin{equation}
  \label{eqn:ann}
  a(t) = \min(|t|, \tau). 
\end{equation}
The cutoff $\tau$ will be chosen large enough to be effectively inactive: 
specifically, hindsight suggests 
\begin{equation}
  \label{eqn:taudef}
  \tau = \max\{|t_{\rm min}-m_{\rm min}r|,|t_{\rm min}-m_{\rm max}r|, |t_{\rm max}-m_{\rm min}r|, |t_{\rm max}-m_{\rm max}r|\}. 
\end{equation}
This 
particular annihilator has been employed in earlier papers on extended 
source inversion 
\cite[]{Plessix:00a,LuoSava:11,Warner:14,HuangSymes:SEG15a,Warner:16,HuangSymes:GEO17}. It
satisfies condition \ref{eqn:abnd} with $C=\lambda$.

Surprisingly, the choice \ref{eqn:ann} yields an optimization
problem with better global convergence properties:

\begin{theorem}
  \label{thm:rampgood}
  For $a(t)=\min\{|t|,\tau\}$ in the definition \ref{eqn:annmult},
  with $\tau$ given by equation \ref{eqn:taudef}, all
  stationary points of $\tJa$ lie in the interval
  \[
    [m_*-\lambda/r, m_*+\lambda/r].
  \]
\end{theorem}
\begin{proof}
  As observed before, $\mbox{supp }\aw[m;d] \subset [t_{\rm
    min}-mr,t_{\rm max}-mr]\subset [-\tau,\tau]$, with $\tau$ defined
  in \ref{eqn:taudef}. Therefore for $t \in  [t_{\rm min}-mr,t_{\rm
    max}-mr]$, $a(t) = |t|$, and reason analogous to the proof of
  Theorem \ref{thm:norminv}, item 2, shows that
  \begin{equation}
    \label{eqn:solnnon}
    \aw[m,d](t)= \left(1+ (4\pi r)^2\alpha^2 t^2\right)^{-1}w_*\left(t+(m-m_*)r\right) 
  \end{equation}
  Straightforward modification of the proof of Theorem
  \ref{thm:epjnonoise} gives
  \begin{equation}
    \label{eqn:expj}
    \tJa[m;d] =\frac{\alpha^2}{2}\int\,dt\,(t-(m-m_*)r)^2(1+(4\pi r)^2 \alpha^2 
    (t-(m-m^*)r)^2)^{-1}w_*(t)^2.
  \end{equation}
  A brief calculation yields
  \begin{equation}
    \label{eqn:gradfinal}
    \frac{\partial}{\partial m}\tJa[m;d] = -r\alpha^2\int \,dt\,\frac{t}{(1+(4\pi r)^2\alpha^2 t^2)^2}w_*(t+(m-m_*)r)^2.
  \end{equation}
  Recall that $w_*(t+(m-m_*)r)$
  vanishes if $|t+(m-m_*)r| > \lambda$. Therefore the integral on the
  RHS of equation \ref{eqn:gradfinal} can be re-written
  \[
    = -r\alpha^2\int_{-(m-m_*)r-\lambda}^{-(m-m_*)r+\lambda}
    \,dt\, \frac{t}{(1+(4\pi r)^2\alpha^2 t^2)^2}w_*\left(t+(m-m_*)r\right)^2
  \]
  If $m > m_*+\lambda/r$, then the entire interval of integration is a proper
  subset of the negative half-axis. Consequently the first factor in the
  integrand satisfies
  \[
    \frac{t}{(1+(4\pi r)^2\alpha^2 t^2)^2}< 0
  \]
  over the interval of integration. The integral is
  negative, so the derivative
  is positive. Similar reasoning applies to the case $m <
  m_*-\lambda/r$. Thus
  \begin{eqnarray}
    m > m_*+\lambda/r & \Rightarrow & \frac{\partial \tJa}{\partial
                                      m}[m;d]> 0; \nonumber\\
    m < m_*-\lambda/r & \Rightarrow & \frac{\partial \tJa}{\partial
                                      m}[m;d]< 0.
                                      \label{eqn:leftright}
  \end{eqnarray}
\end{proof}

\noindent {\bf Remark.} That is, {\em with the choice of penalty
  multiplier $a$ given in equation \ref{eqn:ann}, $\tJa$ has no local minima
  further than  $O(\lambda)$ from the global minimum:} the gradient
has the correct sign and slowness updates computed from it will be
constructive, unless the slowness estimate is already ``within a
wavelength'' of being correct.

\noindent {\bf Remark.} The estimate $|m-m_*|r<\lambda$ for stationary
points $m$ of $\tJa$ is sharp: it is possible to choose $w_* \in \lW$
so that $\lambda - |m-m_*|r$ is as small as you like. In particular,
the ``exact'' or ``true'' slowness $m_*$ is not necessarily the only 
stationary point, or even a stationary point at all.

HERE INSERT RESULT: TJA DIFFBL RIGHTARROW A DIFFBLE - relate in
Discussion to Stolk 03

\section{Extended Inversion}
The preceding theorem established that a proper choice of annihilator
leads to a reduced penalty objective all of whose stationary points
are within $O(\lambda)$ of the target slowness $m_*$, provided that
the data are noise-free in the sense of equation
\ref{eqn:defdatanonoise}. This result leaves open two questions:
\begin{itemize}
\item how does one use this reduced penalty minimization to produce
  a solution of the inverse problem as in problem statement
  \ref{eqn:probstat0}? 
\item how does one answer the same question for noisy data?
\end{itemize}

\subsection{Noise-free Data}
\begin{theorem}
  \label{thm:ipnonoisesuf}
  Suppose that $a$ is given by definition \ref{eqn:ann}, $\alpha$,
  $\mu \in (0,\lambda_{\rm max}]$,
  $d$ is given by
  \ref{eqn:defdatanonoise} with $w_* \in W_{\mu}$, and  $m_{\infty}$ is a stationary
  point of $\tJa[\cdot;d]$. Then $(m_{\infty},\aw[m_{\infty};d])$ is a
  solution of the inverse problem \ref{eqn:probstat0} for any $\lambda
  \ge 2\mu$ and $\epsilon \ge (4\pi r \lambda \alpha)^2$.
\end{theorem}

\begin{proof}
From the
inequalities \ref{eqn:leftright}, $|(m-m_*)r|\le \mu$. From the
identity \ref{eqn:solnnon}, $\mbox{supp }\aw[m_{\infty};d] \subset
[(m_{\infty}-m_*)r-\mu,(m_{\infty}-m_*)r+\mu] \subset
[-2\mu,2\mu]$. Because of the support limitation, $a(t)=|t|$
in the interval of integration appearing in \ref{eqn:residnorm}, so
\[
  e[m_{\infty},\aw[m_{\infty},d];d] 
= 8 \pi^2 r^2 \alpha^4\int^{\mu}_{-\mu}\,dt\,|t-(m_{\infty}-m_*)r|^4(1+(4\pi r)^2 \alpha^2 
|t-(m_{\infty}-m^*)r|^2)^{-2}w_*(t)^2.
\]
and therefore
\begin{equation}
  \label{eqn:estresidnorm}
e[m_{\infty},\aw[m_{\infty};d];d] \le 8 \pi^2 r^2 (2\mu\alpha)^4 \|w_*\|^2 =
\frac{1}{2}(8 \pi r \mu \alpha)^4 \|d\|^2
\end{equation}
\end{proof}

The inequality \ref{eqn:estresidnorm} can be interpreted as a bound 
on $\alpha$, given $\epsilon$ and $\lambda$, for a
stationary point of $\tJa$ to yield a solution of the inverse
problem: one obtains a solution, provided that $\alpha$ is
sufficiently small. On the other hand, it is clear that $\alpha$
cannot be too large if stationary points of $\tJa$ are to yield
solutions: the integrand in \ref{eqn:residnorm} is increasing in
$\alpha$ for every $t$ and $m$, and the multiplier
\[
t \mapsto (4\pi r \alpha(t-(m_{\infty}-m_*)r))^4(1+(4\pi r)^2 \alpha^2 
|t-(m_{\infty}-m^*)r|^2)^{-2}
\]
tends monotonically to $1$ as $\alpha \rightarrow \infty$, uniformly
on the complement of any open interval containing
$t=(m_{\infty}-m_*)r$. Therefore
\begin{equation}
  \label{eqn:elimit}
  \lim_{\alpha \rightarrow \infty}e[m_{\infty},\aw[m_{\infty};d];d] =
  \frac{1}{2}\frac{1}{(4 \pi r)^2}\|w_*\|^2 = \frac{1}{2}\|d\|^2.
\end{equation}
Consequently, there exists $\alpha_{\rm max}(\epsilon,\lambda,d)$ so
that
\[
  e[m_{\infty},\aw[m_{\infty};d];d]  \le \frac{1}{2}\epsilon^2\|d\|^2
  \Rightarrow \alpha \le \alpha_{\rm max}(\epsilon,\lambda,d).
\]
The existence of this limiting penalty weight has been inferred
indirectly; the next section is devoted to a constructive algorithm
for its approximation.

[THE NEXT BIT MAY OR MAY NOT BE WORTH KEEPING - IT WAS INTERESTING TO
WORK OUT]

Without further contraints on data or solution, nothing more can be
said about bounds on $\alpha$. If the data $d$ (and the target wavelet
$w_*$) are assumed to have a square integrable derivative, then a
necessary condition follows from the Heisenberg
inequality (see for example \cite{Folland:07}, p. 255). To formulate
this result in its most general form, introduce the Hilbert subspaces
$V^0 \subset L^2(\bR), V^1 \subset H^1(\bR)$:
\begin{eqnarray}
  V^0 & = & \{ f \in L^2(\bR): Af \in L^2(\bR)\}, \nonumber\\
  V^1 & = & V^0 \cap H^1(\bR).
            \label{eqn:vdef}
\end{eqnarray}
$V^0$ is the domain of $A$, and equipped with the graph norm of $A$. A
natural norm in $V^1$ is
\[
  \|f\|^2_{V^1} = \|f\|_{V^0}^2 + \|f\|_{H^1}^2.
\]
$V^j$ is the completion of $C_0^{\infty}(\bR)$ in the corresponding
norm, j=0,1.

\begin{proposition}
  \label{thm:heis}
For $w \in V^1$,
  \begin{equation}
    \label{eqn:heis}
    \|Aw\|\|w'\| \ge \frac{1}{2}\|w\|^2
  \end{equation}
\end{proposition}

\begin{proof}
  For $w \in C_0^{\infty}(\bR)$,
  \[
    \int w^2 = \left|\int\,dt\, t (w(t)^2)' \right|= \left|2\int\,dt\,tw(t)w'(t)\right| \le
    2\|Aw\| \|w'\|
  \]
  by the Cauchy-Schwarz inequality. Since $C_0^{\infty}(\bR)$ is dense
  in $V^1$, the conclusion follows by continuity.
\end{proof}

In the conventional formulation of the Heisenberg inequality, the $L^2$ norm of
$w'$ is replaced by the its equivalent in terms of the Fourier
transform $\hat{w}$. Adopting temporarily the use of dummy variables
in the expression of functions, the identity \ref{eqn:heis} turns into
the usual form of the Heisenberg inequality: for $w \in V^1$,
\begin{equation}
\label{eqn:fheis}
\|tw(t)\|\|k\hat{w}(k)\| \ge \frac{1}{4\pi}\|w\|^2.
\end{equation}

Define $\krms[w]$, the root mean square estimator of frequency of $w
\in V^1$, by
\begin{equation}
  \label{eqn:krms}
  \krms[w]=\frac{1}{2\pi}\frac{\|w'\|}{\|w\|} = \left(\int
    \,dk\,\frac{|\hat{w}(k)|^2}{\|\hat{w}\|^2} k^2\right)^{1/2}.
\end{equation}
Then the inequalities \ref{eqn:heis}, \ref{eqn:fheis} can be rewritten as
\begin{equation}
  \label{eqn:frmsheis}
  \|Aw\| \ge \frac{\|w\|}{4\pi \krms[w]}.
\end{equation}

For $\lambda >0$, define
\begin{equation}
  \label{eqn:wlam1}
  \lW^1 = \lW \cap H^1(\bR).
\end{equation}
Note that $\lW^1 \subset V^1$ is a closed subspace for any
$\lambda>0$.

\begin{proposition}
  \label{thm:klam}
  For $w \in \lW^1$,
  \[
    \krms[w] \ge \frac{1}{4 \pi \lambda}
  \]
\end{proposition}

\begin{proof}
  Follows directly from inequality \ref{eqn:frmsheis} and the obvious bound
  $\|A|_{\lW}\|\le \lambda$.
\end{proof}
